---
title: "stringr Exercises"
author: "Emma SCHYMANSKI"
date: "2022-10-09"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```

## Background

This little tutorial aims to make you familiar with writing 
RMarkdown while testing some of the functions of the `stringr` 
package and a few regular expressions. Please start by creating 
your own RMarkdown document to write the answers. 

```{r, echo = FALSE}
htmltools::img(src = "https://raw.githubusercontent.com/rstudio/hex-stickers/999a52ccc6b53d8a620a5f930e4f9e68492d346d/PNG/stringr.png",
               style = "float:right;height:100px;")
```

## Part 1: Strings and escape sequences in R

#### 1A. Write a sentence with escape sequences.
Try the sentence: `"It's the end of the world!" he said.\` . 
Assign the string to a variable and try as`print()`, `cat()` and `writeLines()`.

```{r escape sequences}
#write your solutions here
test_string <- '"It\'s the end of the world!" he said.\\'
print(test_string)
cat(test_string)
writeLines(test_string)
```

End of Part 1. 


## Part 2: `stringr` functions 
We will be using the `words` data that is built into `stringr`. 
The data set is available to you if you load the `stringr` package.

```{r words}
length(words)
```

#### 2A. Select words that
1. Contain a `y`

```{r words with y}
#write your solutions here
```


2. Start with `y`
```{r words starting with y}
#write your solutions here
```


3. Contain a `y` within the word

```{r y in word}
#write your solutions here
```


In the next section, we will look into extraction.

#### 2B. Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r extract y and char before}
#write your solutions here
```


Finally, we will look at how to determine string length. 

#### 2C. Get the length of the first ten words

Note: You can use `head(words, 10)` as a convenient way to access the elements of the `words` vector.

```{r length of first 10 words}
#write your solution here
```

End of Part 2. 

In the next section, a little bioinformatics case study will show
how to use string manipulation on viruses 


## Part 3: Virus research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).
For now, just execute the following:

```{r read hepd}
hepd <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

#### 3A. What is the length of the genome sequence?

```{r length of hepd sequence}
#write your solution here
```


#### 3B. What is the sequence composition? How often does each character occur?

```{r sequence composition}
#write your solution here
```

#### 3C. Find motifs in the sequence using `str_locate()`. 

Find all matches of the sequence $ATG$ in the sequence. 

```{r sequence motifs}
#write your solution here
```

